---
title: "R crash course"
output:
  html_document:
    df_print: paged
---

This is an R crash couse, adpated and revised based on an [online tutorial](https://billpetti.github.io/Crash_course_in_R/).

Let's firstly setup a working directory

```{r setup}
knitr::opts_knit$set(root.dir="/home/vmuser/Desktop/PHC6194SPR2021/WK1")
```

We can assign objects using "<-"

```{r}
foo <- "hello world!"
foo
```

R is case-sensitive

```{r error=TRUE}
myvar <- 1
Myvar
myvar
```

We can add comments by using "#"

```{r error=TRUE}
#foo2 <- "hello world!"
foo2
```

Data structures:

Let's look at vector first, which can be defined using c()

```{r}
x <- c(1,2,3,4,5)
x 
firstNames <- c("Shinji","Aska","Rey","Misato")
firstNames
```

We can access the different parts of the vector using index: [idx]

```{r}
firstNames[3]
```

str() will allow us to explore the structure of a vector

```{r}
str(firstNames)
```

R treat categorical variables as factors. We can define that by using as.factor()

```{r}
gender <- c("f","f","f","m","m","m")
gender <- as.factor(gender)
str(gender)
```

List is another data structure in R, which is a sequence of elements of different types

```{r}
myList <- list(x=x,firstNames=firstNames,gender=gender)
myList
```

We can access specific elements within the list

```{r}
myList[[1]]
myList$x
myList[["x"]]
```

Now let's move on to data frames

```{r}
franchise <- c("Mets", "Nationals", "Marlins", "Phillies", "Braves")
city <- c("New York", "Washington, DC", "Miami", "Philadelphia", "Atlanta")
teams <- data.frame(franchise, city)
teams
```

We can get names of variables using colnames()

```{r}
colnames(teams)

```

Let's write our first function: mean calculation

```{r}
our.mean <- function(x){
  return(sum(x) / length(x))
}
```

Let's run the function on some data

```{r}
x <- c(1, 2, 3, 4, 5)
x
our.mean(x) #self-defined mean
```

We can compare it with the built-in mean function

```{r}
mean(x) #built-in mean 
mean(x) == our.mean(x) #check whether these two values are equivalent
```

A more complex function

```{r}
our.summary <- function(x) {
  mean <- mean(x)
  median <- median(x)
  standard_deviation <- sd(x)
  foo <- cbind(mean, median, standard_deviation)
  return(foo)
}

our.summary(x)
```

Install, load, and unload packages

```{r}
install.packages("ggplot2") #install
library(ggplot2) #load
detach("package:ggplot2") #unload
```

Read data into R

```{r}
dat1 <- read.csv("dat1.csv",header = TRUE, na.strings="NA")
str(dat1)
head(dat1)
```

Treat strings as non-factors

```{r}
dat1 <- read.csv("dat1.csv",header = TRUE, na.strings="NA",stringsAsFactors = FALSE)
str(dat1)
head(dat1)

```

We can output data from R

```{r}
write.csv(dat1,"~/Desktop/dat2.csv",row.names = F)
```

Some data manipulation using the built-in data "iris"

```{r}
data(iris) #load the built-in data "iris"
head(iris)
head(iris$Sepal.Length)
```

An easier way to access more than one variables

```{r}
head(with(iris,Sepal.Length/Sepal.Width))

iris$sepal_length_width_ratio <- with(iris, Sepal.Length / Sepal.Width)
head(iris)
iris$sepal_length_width_ratio <- round(iris$sepal_length_width_ratio, 2)
head(iris)
```

Categorize a variable into tertiles

```{r}
iris$sepal_length_width_ratioG<-cut(iris$sepal_length_width_ratio,breaks=quantile(iris$sepal_length_width_ratio,probs=c(0,0.33,0.67,1)),include.lowest = T)
str(iris)
```

Convert factor into numeric

```{r}
iris$sepal_length_width_ratioG<-as.numeric(iris$sepal_length_width_ratioG)
str(iris)
```

Quick summary of a variable

```{r}
summary(iris$sepal_length_width_ratio)

```

Subsetting data

```{r}
unique(iris$Species)

sub_virginica <- subset(iris, Species == "virginica")
head(sub_virginica)
unique(sub_virginica$Species)

ex_virginica <- subset(iris, Species != "virginica")
unique(ex_virginica$Species)

sub_virginica2 <- subset(iris, Species != "virginica" & sepal_length_width_ratio >= 2)
head(sub_virginica2)
```

Select specific variables

```{r}
head(iris[,c(1,3)])
```

Select specific cases/rows

```{r}
iris[c(1:6),]
```

Basic descriptive and summary statistics
```{r}
data(airquality)
summary(airquality)
```

dplyr package

```{r}
if (!require(dplyr)) {
  install.packages(dplyr)
}
```

Frequency tables

```{r}
with(iris,table(Species,Petal.Width))
table(iris$Species,iris$Petal.Width)

with(iris, table(Species, Petal.Width)) %>% prop.table()



```

Column frequencies

```{r}
with(iris, table(Species, Petal.Width)) %>% prop.table(margin = 1)  #row frequencies
```

Row frequencies

```{r}
with(iris, table(Species, Petal.Width)) %>% prop.table(margin = 2)  #column frequencies
cross_column<-with(iris, table(Species, Petal.Width)) %>% prop.table(margin = 2) %>% as.data.frame.matrix() 
```

Basic statistics and modeling

```{r}
dat2<-read.csv("survey_sample_data.csv",header = T, stringsAsFactors = F)
str(dat2)
```

Correlation

```{r}
cor(dat2$Q1,dat2$Q2,use="pairwise.complete.obs")
cor(dat2[,c(2:19)], use = "pairwise.complete.obs")
round(cor(dat2[,c(2:19)], use = "pairwise.complete.obs"),3)
```

Linear regression

```{r}
head(iris)

iris_lm <-lm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width+ factor(Species),data=iris)

summary(iris_lm)

iris_lm$coefficients
confint(iris_lm)
```

Visualization:
Histogram

```{r}
hist(dat2$Q2)

hist(dat2$Q2, main = "Frequency of Responses to Q2", xlab = "Response Value",breaks = c(0.0, 1.0, 2.0, 3.0, 4.0, 5.0))

```

Scatterplots

```{r}
plot(Sepal.Length~Sepal.Width,data=iris)

```

Boxplots

```{r}
boxplot(iris$sepal_length_width_ratio)

```

Use ggplot2 to enhance visualizations

```{r}
if (!require(ggplot2)) {
  install.packages(ggplot2)
}

```

Better scatterplots for iris

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point()

```

Different color by species

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species)) #

```

Differnet size proportional to sepal length width ratio

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio)) #

```

Add some transparency

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) #
```

Add a trend line

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3))+stat_smooth() 
```

Seperate plots by species

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) +stat_smooth() +facet_wrap(~Species) #separate plots by species

```

Density plots

```{r}
if (!require(reshape2)) {
  install.packages(reshape2)
}

dat2_melt<-melt(dat2[,c(2:19)])
ggplot(dat2_melt,aes(value))+geom_density()+facet_wrap(~variable)

```

Boxplots

```{r}
ggplot(iris,aes(y=sepal_length_width_ratio,x=Species))+geom_boxplot()
```

Customize

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) +stat_smooth() +facet_wrap(~Species) 
```

Add title

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) +stat_smooth() +facet_wrap(~Species) + ggtitle("Sepal Length versus Sepal Width") + xlab("Sepal Width")+ylab("Sepal Length") #add title and labels
```

Resize and bold

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) +stat_smooth() +facet_wrap(~Species) + ggtitle("Sepal Length versus Sepal Width") + xlab("Sepal Width")+ylab("Sepal Length") + theme(axis.title=element_text(face="bold",size=14),plot.title=element_text(face="bold",size=16)) #resize and bold
```

Remove the shading

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) +stat_smooth() +facet_wrap(~Species) + ggtitle("Sepal Length versus Sepal Width") + xlab("Sepal Width")+ylab("Sepal Length") + theme(axis.title=element_text(face="bold",size=14),plot.title=element_text(face="bold",size=16))+ theme(panel.background=element_blank()) #remove the shading
```

Use custom color palettes

```{r}
ggplot(iris,aes(x=Sepal.Width,y=Sepal.Length))+geom_point(aes(color=Species,size=sepal_length_width_ratio,alpha=.3)) +stat_smooth() +facet_wrap(~Species) + ggtitle("Sepal Length versus Sepal Width") + xlab("Sepal Width")+ylab("Sepal Length") + theme(axis.title=element_text(face="bold",size=14),plot.title=element_text(face="bold",size=16))+ theme(panel.background=element_blank()) + scale_color_brewer(palette="Greens") #use custom coclr palettes
```

Correlation heatmap

```{r}
dat2_cor_melt<-melt(cor(dat2[,c(2:19)],use="pairwise.complete.obs"))
dat2_cor_melt$value<-round(dat2_cor_melt$value,2)
head(dat2_cor_melt)

ggplot(dat2_cor_melt, aes(Var1, Var2)) + geom_tile(aes(fill = value)) + geom_text(aes(label=value), size = 3, fontface = "bold") + scale_fill_gradient2(low = "#67a9cf", high = "#ef8a62") +theme_minimal() +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  panel.border = element_blank(),panel.background = element_blank(), axis.title = element_blank(), axis.text = element_text(size = 12, face = "bold"))

```

Remove 1s

```{r}
dat2_cor_melt<-dat2_cor_melt[dat2_cor_melt$value!=1,]

ggplot(dat2_cor_melt, aes(Var1, Var2)) + geom_tile(aes(fill = value)) + geom_text(aes(label=value), size = 3, fontface = "bold") + scale_fill_gradient2(low = "#67a9cf", high = "#ef8a62") +theme_minimal() +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  panel.border = element_blank(),panel.background = element_blank(), axis.title = element_blank(), axis.text = element_text(size = 12, face = "bold"))

```

Remove labels

```{r}
ggplot(dat2_cor_melt, aes(Var1, Var2)) + geom_tile(aes(fill = value)) + scale_fill_gradient2(low = "#67a9cf", high = "#ef8a62") +theme_minimal() +theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),  panel.border = element_blank(),panel.background = element_blank(), axis.title = element_blank(), axis.text = element_text(size = 12, face = "bold"))

```

